\paperTitle

### 2014-06-18 doc/nips2013/rebuttal.txt

Thank you for your comments; we will try to respond to your questions below.

# Major Comments

## Regarding Novelty and Section 4 (Examples)

Several reviewers remarked that our development of an algorithm for parameter estimation in latent variable models and the examples we used in Section 4 were derivative from [8,13]. While we certainly build on their work, the contribution of our work is to present an alternative view point to parameter estimation using the method of moments wherein the moments act as constraints for an objective as opposed viewing the problem as one of solving a polynomial. This perspective allows us to open up the space of possible estimation procedures in a very practical sense. We demonstrate this in section 4 by describing how to estimate parameters for various MRF models, models which cannot actually be handled using prior work.

in particular, our presentation of the HMM might be better appreciated if one realizes that it is actually an undirected MRF with general parameter tying. This allows arbitrary coarse features (e.g. all words ending in '-ing') to be included in the model, which has been shown to be extremely effective for several unsupervised language learning tasks in [25]. Further more,
the moment equations for the grid and tall mixture models are considerably complex and it is not clear how the techniques employed in [8,13] could be used to solve them.

% In the HMM example, we were able to use all of the data to get fairly reliable estimates of the emission probabilities ($O$) and consequently avoid several nasty local optimal, as the experiments show.

[25] Painless Unsupervised Learning with Features; Taylor Berg-Kirkpatrick, Alexandre Bouchard-Côté, John DeNero, Dan Klein; NAACL 2010.

## Strong Convexity of Log-Partition Function (Reviewer 3)
If the log-partition function is not strongly convex, standard results show that the model is not identifiable and no procedure can work. Our algorithm will constrain the parameters to the equivalence class of correct parameters and allows regularization, etc. to break ties if desired. In effect, our method does not require the assumption that the log-partition function be strongly convex, but when it is, it finds an exact fit in the exponential family.

## Regarding how the HMMs and mixture models were generated (Reviewer 3)
We generated parameters for both the mixture model and for the
HMMs by generating the model parameter vector $\theta$ with uniform
random entries in [-1,1]. Note that normalization is taken care of by
the partition function.

## Regarding Assumptions 1 and 2 (Reviewer 3)
The assumption that $B$ have full column rank (Assumption 1) is a standard assumption [8,11,13] and is reasonable in practice. From a practical stand-point, if $B$ is rank deficient, then it is indicative that we are over-estimating the number of latent states, and should choose a smaller $k$.
Assumption 2 is certainly strong, but can still be useful when in a large network, we can find a small subset of sources from which we have 3 conditionally independent views (e.g. as in [26])

[26] Unsupervised Learning of Noisy-Or Bayesian Networks; Yoni Halpern, David Sontag; UAI 2013.

## Regarding Section 5 (Reviewer 6)

The size of the composite domain is indeed the product over the domains of the latent variables. With regard to the sensitivity of the unshuffling procedure, the estimates $A_2$ that appear in the regularization term have a sample complexity that is polynomial in the separation between factors; this is not too different from other separation terms that appear in the method of moments algorithms in [8,13].

### 2014-06-18 doc/icml2014/assumption-proof.tex

% Define up front that we will focus on h, x_1, x_2, x_3
In step 1 of \LearnMarginals, we used the bottleneck property of a hidden
  variable $h_i$ to learn conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_v | h_i)$ for every view $x_v \in \sV_{h_i}$ using
  \TensorFactorize. 
In order to do so, we require that \assumptionref{full-rank} holds, i.e.
\begin{assumption*}
  Given a bottleneck $h_1$ with views $x_1, x_2, x_3$, the conditional
  moments $\mOpp{1}{1}, \mOpp{2}{1}, \mOpp{3}{1}$ have full column rank
  $k$, and $\pi\oft 1 \succ 0$.
\end{assumption*}

@@

Without loss of generality, let $i = 1$. We will express the conditional
distribution $\mOpp{v}{1} \eqdef \Pr(x_v | h_1)$ recursively through the
parents of $x_v$, its parents, and so on. The approach is analogous to
message passing, and the key idea is to express the same in the form of
a tensor multiplication. \figureref{message-proof} outlines the
procedure.

@@

Note that $\mYpp{t}{\Pa(h_t)}$ is the conditional probability $\Pr(h_t
 | \Pa(h_t))$, which is assumed to be full rank, and can be easily
 computed from the hidden marginals $Z_\sC$ that contain $h_t$.

>>>>

% Define up front that we will focus on h, x_1, x_2, x_3
In step 1 of \LearnMarginals, we used the bottleneck property of a hidden
  variable $h_i$ to learn conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_v \given h_i)$ for every view $x_v \in \sV_{h_i}$ using
  \TensorFactorize. 
In order to do so, we require that \assumptionref{full-rank} holds, i.e.
\begin{assumption*}
  Given a bottleneck $h_1$ with views $x_1, x_2, x_3$, the conditional
  moments $\mOpp{1}{1}, \mOpp{2}{1}, \mOpp{3}{1}$ have full column rank
  $k$, and $\pi\oft 1 \succ 0$.
\end{assumption*}

@@

Without loss of generality, let $i = 1$. We will express the conditional
distribution $\mOpp{v}{1} \eqdef \Pr(x_v \given h_1)$ recursively through the
parents of $x_v$, its parents, and so on. The approach is analogous to
message passing, and the key idea is to express the same in the form of
a tensor multiplication. \figureref{message-proof} outlines the
procedure.

@@

Note that $\mYpp{t}{\Pa(h_t)}$ is the conditional probability $\Pr(h_t
 \given \Pa(h_t))$, which is assumed to be full rank, and can be easily
 computed from the hidden marginals $Z_\sC$ that contain $h_t$.

### 2014-06-18 doc/icml2014/directed.tex

\begin{figure}
  \centering
  \subfigure[Bottleneck] {
    \includegraphics[width=0.35\columnwidth]{figures/three-view.pdf}
    \label{fig:three-view}
  }
  \hspace{2em}
  \subfigure[Exclusive views]{
    \includegraphics[width=0.35\columnwidth]{figures/exclusive-views.pdf}
    \label{fig:exclusive-views}
  }

@@

%The intuitions of the directed grid generalize readily to general graphs.
%The key property required for our algorithm to work is as follows:
For which subsets of hidden nodes can we recover the marginals?  The following
definition offers a characterization:
\begin{definition}[Exclusive views]
  \label{def:exclusive-views}
Let $S \subseteq \bh$ be a subset of hidden variables.
We say $h_i \in S$ has an exclusive view $x_v$
  if the two conditions hold:
  (i) there exists some observed variable
  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$,
  and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ has full column rank $k$ and can be recovered.
We say that $S$ has the \emph{exclusive views property} if every $h_i \in S$ has an exclusive view.
%If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
\end{definition}

@@

\algorithmref{learnMarginals} summarizes the procedure, \LearnMarginals.
Given $Z_S$, the conditional probability tables for $S$ can easily be
obtained via renormalization.
\begin{theorem}[Hidden marginals from exclusive views]
If $S \subseteq \bx$ is a subset of hidden variables with the exclusive views property,
then \algorithmref{learnMarginals} recovers the marginals $Z_S
= \Pr(\bh_S)$ up to a global relabeling of the hidden variables
determined by the labeling in $\TensorFactorize$.
\end{theorem}

>>>>

\begin{figure}
  \centering
  \subfigure[Bottleneck] {
    \includegraphics[width=0.35\columnwidth]{figures/three-view.pdf}
    \label{fig:bottleneck}
  }
  \hspace{2em}
  \subfigure[Exclusive views]{
    \includegraphics[width=0.35\columnwidth]{figures/exclusive-views.pdf}
    \label{fig:exclusive-views}
  }

@@

%The intuitions of the directed grid generalize readily to general graphs.
%The key property required for our algorithm to work is as follows:
For which subsets of hidden nodes can we recover the marginals?  The following
definition offers a characterization:
\begin{definition}[Exclusive views]
  \label{def:exclusive-views}
Let $S \subseteq \bh$ be a subset of hidden variables.
We say $h_i \in S$ has an exclusive view $x_v$
  if the two conditions hold:
  (i) there exists some observed variable
  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$ (\figureref{exclusive-views}),
  and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ has full column rank $k$ and can be recovered.
We say that $S$ has the \emph{exclusive views property} if every $h_i \in S$ has an exclusive view.
%If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
\end{definition}

@@

\algorithmref{learnMarginals} summarizes the procedure, \LearnMarginals.
Given $Z_S$, the conditional probability tables for $S$ can easily be
obtained via renormalization.
\begin{theorem}[Hidden marginals from exclusive views]
If $S \subseteq \bx$ is a subset of hidden variables with the exclusive views property,
then \algorithmref{learnMarginals} recovers the marginals $Z_S
= \Pr(\bh_S)$ up to a global relabeling of the hidden variables
determined by the labeling from $\TensorFactorize$.
\end{theorem}

### 2014-06-18 doc/icml2014/measurements.tex

\paragraph{Setup}
Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subsets of variables,
and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  \label{eqn:undirectedSetup}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
To some extent, our methods only depend on the conditional independence
structure of the models.

>>>>

\paragraph{Setup}
Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are hidden ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subsets of variables,
and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  \label{eqn:undirectedSetup}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
To some extent, our methods only depend on the conditional independence
structure of the models.

### 2014-06-18 doc/icml2014/tensor-multiplication.tex
Fixing bugs in piecewise
Updates to the pw proof
Updated sigma_max, sigma_min to be sigma_1, sigma_k

\paragraph{Notation}
In the following, let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be
  a $m$-th order tensor and let $\sQ \in \Re^{d'_1 \times \ldots \times
  d'_n}$ be an $n$-th order tensor. %, where wlog, $m \ge n$. 
We use the notation $\Domain(\sP) = [d_1] \times \cdots \times [d_m]$
  to denote the domain of the indices of $\sP$.
Let $M$ be an index matching set, i.e. $M = \{ (a_1; b_1),
  (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresopnding
  dimensions agree, i.e. $d_{a_i} = d'_{b_i}$ for all $i$. Let $D_{\ba}
  = [d_{a_1}] \times \cdots \times [d_{a_\ell}]$ and $D'_{\bb}
  = [d'_{b_1}] \times \cdots \times [d'_{b_\ell}]$ be the domains of the
  corresponding indices.

@@

Note that when $m, n = 2$, i.e. $\sP$ and $\sQ$ are matrices, and $M
  = \{(2,1)\}$, the operation is equivalent to matrix multiplication. 
Furthermore, similar to matrix multiplication, the operator defined above
  is associative and distributive. 
It is also commutative for symmetric tensors.

@@

\begin{theorem}[Tensor Multiplication and Singular Values]
  \label{thm:tensor-multiplication}
Let $\sP, \sQ$ be $m$-th order and $n$-th order tensors respectively, $M$ be an
  index mapping and $\sR = \sP \times_M \sQ$. 
Let $I \subset [m - k + n - k]$ be any arbitrary subset of the index set
  of $\sR$,
and let $I_P$, $I_Q$ be the partitioning of $I$ into subsets contained
  in $\sP$ and $\sQ$ respectively.
Then,
\begin{align*}
\sigma_{\max}(\sR\munf{I}) &\le \sigma_{\max}(\sP\munf{I_P}) \sigma_{\max}(\sQ\munf{I_P}) \\
\sigma_{\min}(\sR\munf{I}) &\ge \sigma_{\min}(\sP\munf{I_Q}) \sigma_{\min}(\sQ\munf{I_Q}).
\end{align*}
\end{theorem}

Note that in the matrix case, when $m = n = 2$ and $M=\{(2,1)\}$, i.e.
when $R = P Q$, we get the standard result that 
\begin{align}
\sigma_{\max}(R) &\le \sigma_{\max}(P) \sigma_{\max}(Q) \\
\sigma_{\min}(R) &\ge \sigma_{\min}(P) \sigma_{\min}(Q) \label{eqn:matrix-singular}.
\end{align}

@@

The key idea is to represent the above expression as matrix multiplication.
To do so, we can temporarily ``inflate'' the unfoldings $\sP\munf{I_P}$
  and $\sQ\munf{I_Q}$ by taking the outer product of $\sP\munf{I_P}$ with
  $|I_Q|$ identity matrices, and vice-versa. Define 
\begin{align*}
  \sP' &\eqdef \sP \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_Q|}} \\
  \sQ' &\eqdef \sQ \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|\bar I_P|}}.
\end{align*}

@@

Using lemma \lemmaref{tensor-prod} and the bound on matrix singular
values, \equationref{matrix-singular}, we get,
\begin{align*}
\sigma_{\max}(\sR\munf{I}) 
    &\le \sigma_{\max}({\sP'}\munf{I'_P}) \sigma_{\max}({\sQ'}\munf{I'_Q}) \\
      &= \sigma_{\max}(\sP\munf{I_P}) \sigma_{\max}(\sQ\munf{I_Q}) \\
    \sigma_{\min}(\sR\munf{I})  
    &\ge \sigma_{\min}({\sP'}\munf{I'_P}) \sigma_{\min}({\sQ'}\munf{I'_Q}) \\
      &= \sigma_{\min}(\sP\munf{I_P}) \sigma_{\min}(\sQ\munf{I_Q}).
\end{align*}
\end{proof}

@@

We will now study the singular values of a tensor projection. 
Let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be a $m$-th order
  tensor. 
For some vector $v \succ 0$, let $\sQ$ be the $m-1$-th order tensor
  formed by projecting one of the axes of $\sP$ by $v$. 
Without loss of generality, let the projected axes be the last one, $\sQ
  \eqdef \sP(\cdot, \cdots, \cdot, v)$.
Then, we have the following lemma linking the singular values of $\sQ$
  and $\sP$.
\begin{lemma}[Tensor projection preserves rank]
  \label{lem:tensor-projection}
  Let $\sP, \sQ$ and $v$ be defined as above. For any unfolding $I
  \not\contains v$, 
\begin{align*}
  \sigma_{\max}(\sQ\munf{I}) &\le \sigma_{\max}(\sP\munf{I}) v_{\max} \\
  \sigma_{\min}(\sQ\munf{I}) &\ge \sigma_{\min}(\sP\munf{I}) v_{\min}.
\end{align*}
In particular, $\rank(\sP\munf{I}) = \rank(\sQ\munf{I})$.
\end{lemma}
\begin{proof}
  We will show that $\sQ\munf{I}$ is $\sP\munf{I}$ related by
  a projection operator. When $v \succ 0$, this projection operator has
  full column rank, entailing the result.

@@

  We can represent this as matrix multiplication by taking the tensor
  product of $v$ with the $D_{\bar I \setminus \{m\}} \times D_{\bar
  I \setminus \{m\}}$ identity matrix, $\bI$. Note that $\bI \otimes v$ is a $
  D_{\bar I \setminus \{m\}} d_m \times D_{\bar I \setminus \{m\}}
  = D_{\bar I } \times D_{\bar I \setminus \{m\}}$ matrix.
  \begin{align*}
    \sQ\munf{I}[\vi;\vj] 
      &= \sum_{j'=\ones, k=1}^{\bd, d_m} \sP\munf{I}[\vi;\vj'\ k] (v[k] \delta[\vj';\vj]) \\
    \sQ\munf{I}
      &= \sP\munf{I} \bI \otimes v.
  \end{align*}
  If $v$ has no non-zero elements, $\bI \otimes v$ has full-column
  rank, with singular values $v_1, \cdots, v_{d_m}$.
\end{proof}

>>>>

\paragraph{Notation}
In the following, let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be
  a $m$-th order tensor and let $\sQ \in \Re^{d'_1 \times \ldots \times
  d'_n}$ be an $n$-th order tensor. %, where wlog, $m \ge n$. 
We use the notation $\Domain(\sP) = [d_1] \times \cdots \times [d_m]$
  to denote the domain of the indices of $\sP$.
Let $M$ be an index matching set, i.e. $M = \{ (a_1; b_1),
  (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresponding
  dimensions agree, i.e. $d_{a_i} = d'_{b_i}$ for all $i$. Let $D_{\ba}
  = [d_{a_1}] \times \cdots \times [d_{a_\ell}]$ and $D'_{\bb}
  = [d'_{b_1}] \times \cdots \times [d'_{b_\ell}]$ be the domains of the
  corresponding indices.

@@

Note that when $m, n = 2$, i.e. $\sP$ and $\sQ$ are matrices, and $M
  = \{(2,1)\}$, the operation is equivalent to matrix multiplication. 
Furthermore, similar to matrix multiplication, the operator defined above
  is associative and distributive.

@@

\begin{theorem}[Tensor Multiplication and Singular Values]
  \label{thm:tensor-multiplication}
Let $\sP, \sQ$ be $m$-th order and $n$-th order tensors respectively, $M$ be an
  index mapping and $\sR = \sP \times_M \sQ$. 
Let $I \subset [m - k + n - k]$ be any arbitrary subset of the index set
  of $\sR$,
and let $I_P$, $I_Q$ be the partitioning of $I$ into subsets contained
  in $\sP$ and $\sQ$ respectively.
Then,
\begin{align*}
\sigma_{1}(\sR\munf{I}) &\le \sigma_{1}(\sP\munf{I_P}) \sigma_{1}(\sQ\munf{I_P}) \\
\sigma_{k}(\sR\munf{I}) &\ge \sigma_{k}(\sP\munf{I_Q}) \sigma_{k}(\sQ\munf{I_Q}).
\end{align*}
\end{theorem}

Note that in the matrix case, when $m = n = 2$ and $M=\{(2,1)\}$, i.e.
when $R = P Q$, we get the standard result that 
\begin{align}
\sigma_{1}(R) &\le \sigma_{1}(P) \sigma_{1}(Q) \\
\sigma_{k}(R) &\ge \sigma_{k}(P) \sigma_{k}(Q) \label{eqn:matrix-singular}.
\end{align}

@@

The key idea is to represent the above expression as matrix multiplication.
To do so, we can temporarily ``inflate'' the unfoldings $\sP\munf{I_P}$
  and $\sQ\munf{I_Q}$ by taking the outer product of $\sP\munf{I_P}$ with
  $|I_Q|$ identity matrices, and vice versa. Define 
\begin{align*}
  \sP' &\eqdef \sP \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|I_Q|}} \\
  \sQ' &\eqdef \sQ \otimes I_{d'_1} \otimes \cdots \otimes I_{d'_{|\bar I_P|}}.
\end{align*}

@@

Using lemma \lemmaref{tensor-prod} and the bound on matrix singular
values, \equationref{matrix-singular}, we get,
\begin{align*}
\sigma_{1}(\sR\munf{I}) 
    &\le \sigma_{1}({\sP'}\munf{I'_P}) \sigma_{1}({\sQ'}\munf{I'_Q}) \\
      &= \sigma_{1}(\sP\munf{I_P}) \sigma_{1}(\sQ\munf{I_Q}) \\
    \sigma_{k}(\sR\munf{I})  
    &\ge \sigma_{k}({\sP'}\munf{I'_P}) \sigma_{k}({\sQ'}\munf{I'_Q}) \\
      &= \sigma_{k}(\sP\munf{I_P}) \sigma_{k}(\sQ\munf{I_Q}).
\end{align*}
\end{proof}

@@

We will now study the singular values of a tensor projection. 
Let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be a $m$-th order
  tensor. 
For some vector $v \succ 0$, let $\sQ$ be the $m-1$-th order tensor
  formed by projecting one of the axes of $\sP$ by $v$. 
Without loss of generality, let the projected axes be the last one, $\sQ
  \eqdef \sP(\cdot, \cdots, \cdot, v)$.
Then, we have the following lemma linking the singular values of $\sQ$
  and $\sP$.
\begin{lemma}[Tensor projection preserves rank]
  \label{lem:tensor-projection}
  Let $\sP, \sQ$ and $v$ be defined as above. For any unfolding $I
  \not\contains v$, 
\begin{align*}
  \sigma_{k}(\sP\munf{I}) &\ge \frac{\sigma_{k}(\sQ\munf{I})}{v_{\max}}.
\end{align*}
In particular, if $\sQ\munf{I}$ has full row/column rank, $\sR\munf{I}$
  does too.
\end{lemma}
\begin{proof}
  We will show that $\sQ\munf{I}$ is $\sP\munf{I}$ related by
  a projection operator. When $v \succ 0$, this projection operator has
  full column rank, entailing the result.

@@

  We can represent this as matrix multiplication by taking the tensor
  product of $v$ with the $D_{\bar I \setminus \{m\}} \times D_{\bar
  I \setminus \{m\}}$ identity matrix, $\bI$. Note that $\bI \otimes v$ is a $
  D_{\bar I \setminus \{m\}} d_m \times D_{\bar I \setminus \{m\}}
  = D_{\bar I } \times D_{\bar I \setminus \{m\}}$ matrix.
  \begin{align*}
    \sQ\munf{I}[\vi;\vj] 
    &= \sum_{j'=\ones, k=1}^{\bd, d_m} \sP\munf{I}[\vi;\vj'\ k] \underbrace{(v[k] \delta[\vj' \vj])}_{v'[k \vj'; \vj]} \\
    \sQ\munf{I}
      &= \sP\munf{I} \bI \otimes v.
  \end{align*}
  If $v$ has no non-zero elements, $\bI \otimes v$ has full-column
  rank, with singular values $v_1, \cdots, v_{d_m}$. 
  Matrix multiplication can not increase the rank of the matrix, and
  thus $\rank(\sP\munf{I}) \ge \rank(\sQ\munf{I})$.

\end{proof}
